Working title: the most amaz-zing SPIM and how we built it.

### 2015-01-15 latex-style-files-spie-manuscript/manuscript1.0.tex
-deleted OOP paragraph -rewritten software into
-last commit before splitting between ludo and caroline

@@

The microscope was designed and developed by optical technology specialists yet needs to be operational in a multi-user environment. The greatest challenge consists in making the use of the microscope easy and intuitive such that no formidable expertise is needed when it is operated by interdisciplinary researchers. Another aspect of software design is avoiding redundancy in writing software, for example, when a second or third microscope with basic similarity yet slightly specialised equipment is being designed in the same laboratory. Writing well structured scientific software in a group of researchers is an often recurring challenge. The final code should be the common effort of all contributors each with their own know-how and practical expertise. Contemporary experimental research often relies on the acquisition and real-time processing of various data streams. In a given experiment, data may be acquired from multiple, even spatially separated sources, while individual processing steps may be computationally intensive and need to be distributed over several computers to be performed on-line. We chose to write our software in LabVIEW (National Instruments), as this language is particularly well suited to hardware control and data acquisition by pre-developed methods, using the Murmex library. The Murmex library, developed by one of the co-authors, helps users to collaborate on software development by letting them independently create their respective contributions to the software within a standardized programming scheme by employing object oriented programming methodology (OOP). Murmex uses the library LabbitMQ, which is a wrapper of RabbitMQ, a message-oriented middle ware facilitating message shipment and receival. These layers of abstraction are mostly hidden for the developers.

\subsection{Object oriented programming}
In OOP the developer creates classes, which are data structures that contain private data and methods that perform actions on the data. LabVIEW supports classes since 2006, but got to its mature form with LabVIEW 2009. When you create a subclass (child) from the toplevel class (parent), you automatically inherit its methods and data. Public methods can be called by the child, while private methods are only accessible by the owning class. A method implemented by the parent, which has the Dynamic Dispatch enabled and is public, can be replaced or augmented by the child. The parent can force the child to call its method or allow to ignore its code. A call to such a dynamic and public method is chained, it thus calls first the child method, which perform the action before or after calling the parent. The parent calls again its parent before or after executing its code and thus create a complete call chain for a single method.
Next to the methods, the classes have private class data, which can only be accessed by the owning class. In order to give access to its data, a method needs to have methods accessing the data. The instance created at runtime from a class is called the object. Many objects can be made from a class, like many pies can be made from a pie recipe. Next, each object runs independently and contains its own data.

\subsection{RabbitMQ}
The RabbitMQ server (message broker) needs to be installed either locally or somewhere in the network. The message-based communication allows for buffered, asynchronous and reliable message communication between several software components. Messages are routed by the broker to the specific software component by its ID. In order to access the functionality of RabbitMQ, the application programming interface (API) \href{http://sine.ni.com/nips/cds/view/p/lang/en/nid/211065}{LabbitMQ} was used, which uses .NET calls for sending and receiving messages to and from the RabbitMQ broker. Messages can be sent to a single recipient by its ID or sent to all available components.

>>>>

\section{Optical characterisation}

@@

The microscope was designed and developed by optical technology specialists yet needs to be operational in a multi-user environment. The greatest challenge therefore consists in making the use of the microscope easy and intuitive such that no formidable expertise is needed when it is operated by interdisciplinary researchers. Our experimental research is based on the acquisition and real-time processing of various data streams. In a given experiment, data may be acquired from multiple, spatially separated sources, and individual processing steps can be computationally intensive and may need to be distributed over several computers to be performed at run-time. For those reasons we chose to write our software in LabVIEW (National Instruments), as this language is particularly well suited for hardware control and data acquisition through pre-developed methods.

Another important aspect of software design is avoiding redundancy when scaling up the software, for example, when a second or third microscope with basic similarity yet slightly specialised equipment is being designed in the same laboratory. Writing well structured, scalable and easily maintainable scientific software in a group of researchers is an often recurring challenge. Object oriented programming is a programming paradigm whose main aspects are encapsulation, inheritance and polymorphism and a useful tool to address these challenges.

Encapsulation refers to the concept that the entirety of software needed to say operate a microscope is broken up into self-contained modules. Those modules are user-defined data types and referred to as ``classes" where one instance of a class is called an ``object". For example a class could be the basic code to operate an  abstract laser while 5 instances of this class would be the necessary adjustments to operate 5 real world lasers from different manufacturers. This modular programming style keeps large amounts of software routines separate and less prone to conflict with each other.

Classes are arranged in hierarchical structures and inheritance, the act of passing on knowledge from a ``parent" to a ``child" within a class, reduces the amount of programming required when increasingly more complex functions have to be added to the system. For example when specialised behaviour is added to the bottom of the hierarchy, only code regarding this very specific task has to be added while every other aspect is inherited from above. For example, if the code for a motor stage has already been implemented, a rotation stage can be interpreted as the child of a translation stage where the travel range is infinite. This increases code re-usability and saves time.

In this way, the microscope software is decomposed into the communication and interaction of different objects through methods, bits of code, which act on each object's private data.

The Murmex software development kit is based on the OOP paradigm and helps users to collaborate on software development by letting them independently create their respective contributions to the software in the form of new objects or methods within a standardized programming scheme.

\subsection{RabbitMQ}
Murmex uses the library LabbitMQ, which is a wrapper of RabbitMQ, a message-oriented middle ware facilitating message shipment and receival. These layers of abstraction are mostly hidden for the developers.
The RabbitMQ server (message broker) needs to be installed either locally or somewhere in the network. The message-based communication allows for buffered, asynchronous and reliable message communication between several software components. Messages are routed by the broker to the specific software component by its ID. In order to access the functionality of RabbitMQ, the application programming interface (API) \href{http://sine.ni.com/nips/cds/view/p/lang/en/nid/211065}{LabbitMQ} was used, which uses .NET calls for sending and receiving messages to and from the RabbitMQ broker. Messages can be sent to a single recipient by its ID or sent to all available components.
